---
title: "TnSeq_practice"
author: "Alena Marstul"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
Lets load the libraries we are going to use in this practice section
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(readr)
```

Lets load the file with experimental results from TnSeq experiment

The path to the file is different for you. Please make sure to specify a correct path. Let Alena know if you have any problems with that.
```{r}
#Note: tbl_df function forwards the argument to "as_data_frame""

exp_data<-read.csv("~/Documents/1st_Year/TAing/BIMM143/BIMM143_data/Experimental_results.csv")
head(exp_data)
```
As you see, this table has several columns. Column "Strain" shows which strain was used in the experiment. "Environment" shows conditions in which microorganisms were propagated. Columns "Mut_ID" and "BC_ID" give information about type of mutations and barcodes that correspond to them. Columns "H_0": "H_96" show frequenfy of each barcode over time.

Lets modify a table:

First, remove the column "X". We are not going to use the information contained in it.
For this we are going to use function "select"
```{r}
# Remove column "X" from a dataframe:
exp_data<-select(exp_data, -X)
head(exp_data)
```
Lets get ready for plotting:

We are going to use ggplot2 package. Please check the syntax that is commonly used for plotting with this package. Does everything make sense? 

```{r}
# Information from a ggplot description page:
# ggplot(df, aes(x, y, <other aesthetics>))
```

To make a graph, we need two give a function 2 variables (2 columns) to plot them against each other.

What variables are we going to use?

Lets rearrange our table to be able to plot the data easily.
Instead on keeping information about barcode frequency in rows, we are going to create a column "Time" with time points and a column "Frequency" with corresponding barcode frequencies.

* Explanation on the board.

```{r}
# First, check how function "gather" works
exp_rearranged<-gather(exp_data, Generation, Frequency,H_0:H_96)
head(exp_rearranged)
```
You might have noticed that "Generation" column contains both "H" that stands for "hours" and numbers. Lets remove "H_" part from this column.

Check the syntax of "separate" function.
```{r}
# Separate values in "Generation" column into 2 columns
table_for_graph<-separate(exp_rearranged,Generation,into=c("H","Time"))
head(table_for_graph)

# Remove column "H" using function "select"
table_for_graph<-select(table_for_graph, -H)
table_for_graph
```
You might have noticed that our table contains a lot of "NA" values.
Go ahead and remove them with na.omit function. Don't forget to check it's syntax first!
```{r}
table_cleaned<-na.omit(table_for_graph)
table_cleaned$Time<-as.numeric(table_cleaned$Time)
head(table_cleaned)
```
Now the table is ready.
How are we going to plot all the values? Do we need to separate them in any way? If yes, then how?

```{r}
# We might need to plot data for each strain separately..
DivAnc<-filter(table_cleaned, table_cleaned$Strain=="DivAnc")
L013<-filter(table_cleaned, table_cleaned$Strain=="L013")


```

```{r}
# make a plot for DivAnc strain
DivAnc_plot=ggplot(DivAnc)+geom_line(aes(x=Time,y=Frequency,group=BC_ID),alpha=.2,colour="#000033")+ggtitle("DivAnc_SC3")+theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")
DivAnc_plot
```

```{r}
# make a plot for L013 strain
L013_plot=ggplot(L013)+geom_line(aes(x=Time,y=Frequency,group=BC_ID),alpha=.2,colour="#CC6633")+ggtitle("L013_SC3")+theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")
L013_plot
```


Can we make 2 graphs at the same time?

```{r}
ggplot(table_cleaned)+geom_line(aes(x=Time,y=Frequency,group=BC_ID),alpha=.2,colour="#000033")+facet_grid(.~Strain)+ggtitle("Barcode trajectories")+theme(plot.title = element_text(hjust = 0.5))+xlab("Time, hours") + ylab("Log10(Barcode frequency)")

```

Lets pick one mutation and check how it behaves in different strains

```{r}
# I've chosen Mut_ID==34
mut34<-filter(table_cleaned, table_cleaned$Mut_ID=="34")
mut34          
```
```{r}
ggplot(mut34,aes(Time, Frequency, group=BC_ID, color=BC_ID))+geom_line()+theme(legend.position="none")+facet_grid(.~Strain)+ggtitle("Mutation_34")+xlab("Time, hours") + ylab("Log10(Barcode frequency)")+theme(plot.title = element_text(hjust = 0.5))
```
What can you tell looking at this plot?
Why do we have 2 clusters of barcodes?

Lets filter out barcodes with frequency > (-5) and use them for subsequent analysis.
```{r}
mut34_f<-filter(mut34, mut34$Frequency>(-5))
mut34_f
```

Plot again the same type of graph, but use filtered data. Make sure that you have done everything right.

```{r}
ggplot(mut34_f,aes(Time, Frequency, group=BC_ID, color=BC_ID))+geom_line()+theme(legend.position="none")+facet_grid(.~Strain)+ggtitle("Mutation_34")+xlab("Time, hours") + ylab("Log10(Barcode frequency)")+theme(plot.title = element_text(hjust = 0.5))
```


```{r}
ggplot(mut34_f,aes(Time, Frequency, colour = BC_ID, group=BC_ID))+geom_point()+geom_smooth(se = FALSE, method = "lm")+facet_grid(.~Strain)+theme(legend.position="none")+ggtitle(paste("Mutation",34, sep="_"))+xlab("Time, hours")+ ylab("Log10(Barcode frequency)")
```
Now you chan choose a different mutation and check how it behaves across strains.

Now it's time to estimate slope for each barcode.
Lets greate a file that will contain information about BC_ID, Mut_ID, Strain, and estimated slope.

```{r}
# Lets become familiar with lm function:

# For this exercise, take the filtered data for mutation 34 (mut34_f) and filter out information about one barcode you like

# I have chosen BC_ID=25361 in DivAnc strain
BC_25361<-filter(mut34_f, mut34_f$BC_ID=="25361", mut34_f$Strain=="DivAnc")
BC_25361
```
```{r}
#Lets plot frequency of this barcode:
ggplot(BC_25361,aes(Time, Frequency, colour = BC_ID))+geom_point()+theme(legend.position="none")+ggtitle("BC_25361")+xlab("Time, hours") + ylab("Log10(Frequency)")
```
```{r}
#Lets use lm function to fit the line to these points:
ggplot(BC_25361,aes(Time, Frequency, colour = BC_ID))+geom_point()+geom_smooth(se = FALSE, method = "lm")+theme(legend.position="none")+ggtitle("BC_25361")+xlab("Time, hours") + ylab("Log10(Frequency)")
```

```{r}
# Lets check what data does lm function return:
regression_model<-lm(Frequency~Time,BC_25361)
summary_data<-summary(regression_model)
summary_data
```
```{r}
# The information we are interested in is the value of Slopeand Intercept of this line:
# Let's try to access them:

# Time
Time<-summary_data$coefficients[2]
Time

# Intercept:
Intercept<-summary_data$coefficients[1]
Intercept

```
 Now we can find slopes for each barcode for each mutation in all strains.

```{r}
# Lets create the file:
data_header=matrix(data = NA,nrow = 1,ncol = 7)
        data_header[1]="Mut_ID"
        data_header[2]="BC_ID"
        data_header[3]="Strain"
        data_header[4]="Slope"
        data_header[5]="Intercept"
        data_header[6]="R^2"
write.table(data_header,"~/Documents/1st_Year/TAing/BIMM143/BIMM143_data/Tnseq_practice_output.csv",append = FALSE, sep = ",",eol="\n",dec=".",row.names = FALSE,col.names = FALSE)
```


```{r}
for (mut in unique(table_cleaned$Mut_ID)) {
    mut_data=filter(table_cleaned,table_cleaned$Mut_ID==paste(mut))
    #now we have all data for each mutation separately
    for (bc in unique (mut_data$BC_ID)) {
      #now we filtered data for each barcode within 1 mutation
      bc_mut_data=filter(mut_data,mut_data$BC_ID==paste(bc))
      for (strain in unique (bc_mut_data$Strain)) {
        str_bc_mut_data=filter(bc_mut_data,bc_mut_data$Strain==paste(strain))
        #only considering combinations with 3 or more data points - anything less is statistically insignificant
        if (nrow(str_bc_mut_data)>2){
          regression_model=lm(Frequency~Time,str_bc_mut_data)
          summary_data=summary(regression_model)
          #now write to the output file! Prepare the data array first
          data_output=matrix(data = NA,nrow = 1,ncol = 6)
          data_output[1]=mut
          data_output[2]=bc
          data_output[3]=strain
          #slope
          data_output[4]=summary_data$coefficients[2]
          #intercept
          data_output[5]=summary_data$coefficients[1]
          #r-squared
          data_output[6]=summary_data$r.squared
          #time to write
          write.table(data_output,"~/Documents/1st_Year/TAing/BIMM143/BIMM143_data/Tnseq_practice_output.csv",append = TRUE, sep = ",",eol="\n",dec=".",row.names = FALSE,col.names = FALSE)
      }
    }
  }
 }

  


```

